Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-15 ieee_paper.tex
reformating code

The values from the non zero element in the matrix are the simplest to
load. One can load the 16 floating point values for the 4 non zero of
the matrix at once. The 4 floats coming from the four matrices for
each non zero are naturally grouped within each lane of the SIMD
vector. One can use lane permutation to broad a lane to take the whole
vector. So one 1 permutations and a quarter load are necessary to put
the data of the matrix in the required format.

@@

\begin{figure*}
\begin{verbatim}
__m512 read_aaaa(float* a)                                                              
{
    int int_mask = 0x1111;
    __mmask16 mask = _mm512_int2mask(int_mask);
    __m512 v1;
    v1 = _mm512_setzero_ps();
    v1 = _mm512_mask_loadunpacklo_ps(v1, mask, a);
    v1 = _mm512_mask_loadunpackhi_ps(v1, mask, a);
    v1 = _mm512_swizzle_ps(v1, _MM_SWIZ_REG_AAAA);
    return v1_old;
}
\end{verbatim}
\caption{read\_aaaa code}
\label{code:read_aaaa}
\end{figure*}

Figure~\ref{code:permute} is a function that takes four floats from
float register \ttt{v1}, and places them in each of the four
lanes. Because the permulation function only operates on 4-byte
integers, it is necessary to convert (in place) each float to an
integer (the bit structure is not changed), permute the lanes and cast
the integers back to floats (with no modification of the bit
structure)

\begin{figure*}
\begin{verbatim}
__m512 permute(__m512 v1, _MM_PERM_ENUM perm)                                           
{
    __m512i vi = _mm512_castps_si512(v1);
    vi = _mm512_permute4f128_epi32(vi, perm);
    v1 = _mm512_castsi512_ps(vi);
    return v1;
}
\end{verbatim}
\caption{permute code}
\label{code:permute}
\end{figure*}

>>>>

The values from the non zero element in the matrix are the simplest to
load. One can load the 16 floating point values for the 4 non zero of
the matrix at once. The 4 floats coming from the four matrices for
each non zero are naturally grouped within each lane of the SIMD
vector. One can use a swizzle operation to broadcast the 4 elements of
a single matrix to take the whole vector.

Loading the entries from the {\tt x} vector is a little bit more
complicated.

@@

\begin{figure*}
\begin{verbatim}
const __m512i offsets = _mm512_set4_epi32(3,2,1,0);
const __m512i four = _mm512_set4_epi32(4,4,4,4); 
int int_mask = 0x1111;
__mmask16 mask = _mm512_int2mask(int_mask);

#pragma omp parallel for
for (int r=0; r < nb_rows; r++) {
  __m512 accu = _mm512_setzero_ps();

  for (int n=0; n < nz; n+=4) {
    float* a = &dom.col_id_t[0]+n+nz*r;
    __m512i vect_i = _mm512_setzero_ps();
    vect_i = _mm512_mask_loadunpacklo_ps(vect_i, mask, a);
    vect_i = _mm512_mask_loadunpackhi_ps(vect_i, mask, a);
    vect_i = _mm512_swizzle_ps(vect_i, _MM_SWIZ_REG_AAAA);
    vect_i = _mm512_fmadd_epi32(vect_i, four, offsets);

    all_vect = _mm512_i32gather_ps(vect_i, dom.vec_vt, scale);

    mat_ent = _mm512_load_ps(dom.data_t + nb_mat*(n + r*nz));

    //perform first tensor multiplication
    correct_vect = permute(all_vect, _MM_PERM_AAAA);
    correct_mat = _mm512_swizzle_ps(mat_ent, _MM_SWIZ_REG_AAAA);
    accu = _mm512_fmadd_ps(correct_vect, correct_mat, accu);

    //three more times with BBBB, CCCC, and DDDD
  }
  _mm512_storenrngo_ps(dom.result_vt+nb_mat*nb_vec*r, accu);  
} 
\end{verbatim}
\caption{mat mul code}
\label{code:mat_mul}
\end{figure*}

Note that the ``permute'' function is using the {\tt
  \_mm512\_permute4f128\_epi32} instruction and therefore typecast are
necessary to convert to a vector of int type.
